---
title: "CPTS575_Homework2_Zain_Mudassar(011712157)"
author: "Zain Mudassar"
date: "9/11/2020"
output: pdf_document
---

#Question 1:

(a) Use the read.csv() function to read the data into R, or the csv library to read in the
data with python. In R you will load the data into a dataframe. In python you may store it as a list
of lists or use the pandas dataframe to store your data. Call the loaded data college. Ensure that
your column headers are not treated as a row of data.

```{r, Part a}



college <- read.csv(file = "data/College.csv");
head(college)

```

(b) Find the median cost of books for all schools in this dataset.

```{r, Partb}

median(college$Books)
```


(c) Produce a scatterplot that shows a relationship between two numeric (not factor or
boolean) features of your choice in the dataset. Ensure it has appropriate axis labels and a title.

```{r, Partc}

plot(college$Apps,college$Accept, main="Applications Vs Acceptance", xlab="Applications ", ylab="Accepted ", pch=19 ,
     col = c("green","red"))

lines(smooth.spline(college$Apps,college$Accept),lwd=3)

```

(d) Produce a histogram showing the overall enrollment numbers (P.Undergrad plus
F.Undergrad) for both public and private (Private) schools. You may choose to show both on a
single plot (using side by side bars) or produce one plot for public schools and one for private
schools. Ensure whatever figures you produce have appropriate axis labels and a title.

```{r, Partd}
priCollege <- college[college$Private == "Yes",];
pubCollege <- college[college$Private == "No",];
 
hist(priCollege$F.Undergrad+priCollege$P.Undergrad,xlab="Private Schools",ylab = "Total Enrollments",main = "Histogram of Private College")

hist(pubCollege$F.Undergrad+pubCollege$P.Undergrad,xlab="Public Schools",ylab = "Total Enrollments",main = "Histogram of Public College")

```

(e) Create a new qualitative variable, called Top, by binning the Top10perc variable into
two categories (Yes and No). Specifically, divide the schools into two groups based on whether
or not the proportion of students coming from the top 10% of their high school classes exceeds
75%.
Now produce side-by-side boxplots of the schools' acceptance rates (based on Accept and Apps)
for each of the two Top categories. There should be two boxes on your figure, one for top
schools and one for others. How many top universities are there?

```{r, Parte}
for (i in 1: nrow(college)) {
  
  if(college$Top10perc[i]<75)
  {college$Top[i] <- "No"
  }
 else if(college$Top10perc[i]>75)
  {
    college$Top[i] <- "Yes"} 
}


boxplot(college$Apps/college$Accept ~college$Top  ,  col = "orange" , main="Acceptance rate of Top",xlab = "Top" ,ylab="Acceptance")

#Number of Top univerities 

result<-"Total number of Top universities are"
result
length(which(college$Top=="Yes"));





```

(f) Continue exploring the data, producing two new plots of any type, and provide a brief
(one to two sentence) summary of your hypotheses and what you discover. Feel free to think
outside the box on this one but if you want something to point you in the right direction, look at
the summary statistics for various features, and think about what they tell you. Perhaps try
plotting various features from the dataset against each other and see if any patterns emerge.

```{r, Partf}

plot(x = college$Apps,
y = college$Accept ,
xlab ="Apps",
ylab = "Accept",
main = "Applications and Acceptance", pch=19, 
col = c("blue","red"))
lines(smooth.spline(college$Apps,college$Accept),lwd=3)


plot(x = college$F.Undergrad,
y = college$Accept ,
xlab ="F.Undergrad",
ylab = "Accept",
main = "F.Undergrad Vs Accept",pch=19 ,
col = c("blue","red"))
lines(smooth.spline(college$F.Undergrad,college$Accept),lwd=3)

plot(x = college$F.Undergrad,
y = college$Enroll ,
xlab = "F.Undergrad",
ylab = "enroll",
main ="F.undergrad vs Enroll",pch=19 ,
col = c("blue","red"))
lines(smooth.spline(college$F.Undergrad,college$Enroll),lwd=3)

```

Summary: 

Based on information in the graphs we can comment on the relationship between  of the variables  such as number of applications and Acceptance rate have a linear relationship. With the increase in the number of applications, the acceptance rate also increases.


Secondly, total full time undergraduates in the college and Accept also have an almost linear relationship. Number of Full time undergraduates with the increase in acceptance rate. 

Lastly, F.Undergrad and Enroll also have almost a linear relationship between them. When enrollment increases there will also be a increase in F.Undergrad. 




##Question 2

(a) Specify which of the predictors are quantitative (DEFINE), and which are qualitative
(DEFINE)? Keep in mind that a qualitative variable may be represented as a quantitative type in
the dataset, or the reverse. You may wish to adjust the types of your variables based on your
findings.

```{r, Q2Parta}
auto <- read.csv(file = "data/Auto.csv")
head(auto)


       
auto[auto=="?"] <-NA
auto<-
  auto[complete.cases(auto),]






quantitative<-c("mpg","cylinders" , "displacement" ,"horsepower" ,"weight" ,"acceleration")

qualitative<-c("year" , "origin" , "name")
quantitative
qualitative
```

(b) What is the range, mean and standard deviation of each quantitative predictor?

```{r, Q2Partb}
sapply((list(auto$mpg, auto$cylinders, auto$displacement,auto$horsepower, auto$weight,auto$acceleration)),range)

sapply((list(auto$mpg, auto$cylinders, auto$displacement, auto$weight,auto$acceleration,as.numeric(auto$horsepower))),mean)

sapply((list(auto$mpg, auto$cylinders, auto$displacement, auto$weight,auto$acceleration,as.numeric(auto$horsepower))),sd)

```

(c) Now remove the 40th through 80th (inclusive) observations from the dataset. What is
the range, mean, and standard deviation of each predictor in the subset of the data that remains?

```{r, Q2Partc}
 excluded<- auto[-c(40:80), ]
excluded
sapply((list(excluded$mpg, excluded$cylinders,excluded$displacement,excluded$horsepower, excluded$weight,
             excluded$acceleration)), range)

sapply((list(excluded$mpg, excluded$cylinders,excluded$displacement,excluded$weight,excluded$acceleration)), mean)
sapply((list(excluded$mpg, excluded$cylinders,excluded$displacement, excluded$weight, excluded$acceleration)), sd)

```

(d) Using the full data set, investigate the predictors graphically, using scatterplots,
correlation scores or other tools of your choice. Create a correlation matrix for the relevant
variables.

```{r, Q2Partd}

par(mfrow=c(2,2))
plot(x = auto$horsepower,
     y = auto$mpg,
     xlab = "horsepower",
     ylab = "mpg",
     main = "Horse Power vs MPG",
     col = c("green","red"))


plot(x = auto$cylinders,
     y = auto$mpg,
     xlab = "cylinders",
     ylab = "MPG",
     main = "Cylinders vs MPG",
     col = c("green","red"))

plot(x = auto$displacement,
     y = auto$mpg,
     xlab = "Displacement",
     ylab = "MPG",
     main = "Displacement vs MPG",
     col = c("green","red"))

plot(x = auto$weight,
     y = auto$mpg,
     xlab = "Weight",
     ylab = "MPG",
     main = "Weight vs MPG",
     col = c("green","red"))

cor.test(auto$mpg,auto$acceleration)
cor.test(auto$mpg,auto$weight)
cor.test(auto$mpg,auto$cylinders)
cor.test(auto$mpg,auto$displacement)
```

(e) Suppose that we wish to predict gas mileage (mpg) on the basis of the other variables.
Which, if any, of the other variables might be useful in predicting mpg? Justify your answer
based on the prior correlations

ANSWER: The prior correlation data show that there is a moderate positive correlation between 'mpg' and 'acceleration' of 0.42 which means that as acceleration takes more time, mpg also increases. The second correlation is between 'mpg' and 'weight' which shows a strong negative correlation of 0.83 between two variables, meaning that if weight increases the mpg decreases.
Similarly, there is a strong negative correlation of 0.80 between 'mpg' and 'displacement' which means that as displacement increases mpg decreases.
Lastly, we made a correlation between 'mpg' and 'cylinders' and the result shows that there is a negative correlation of 0.77 between two variables, meaning if cylinders increase mpg decreases.
Results:If displacement, cylinders and weight increase then mpg decreases and if acceleration takes more time than there is a moderate increase in mpg.
